#Lab_03 {lab_03}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)

# Load janitor here
library(janitor)

```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)

# your answer here

Maryland <- rio::import("https://docs.google.com/spreadsheets/d/1-7VTVoQs_YnD-6eSY3G3rKbYThiceIfLOus4skF_l1c/edit#gid=1282231278", which = "Maryland")

```


## Answer questions

Q1: Which five cities have the largest overall Hispanic population in 2020? Name the five cities that had the largest Hispanic population in 2020. 

A1:
```{r}

# Write your code in this codeblock.

Hispanic <- subset(Maryland, select = c(hispanic_2020))
summary(Hispanic$hispanic_2020)


Maryland %>% 
     select(hispanic_2020, place) %>% 
     slice_max(hispanic_2020, n = 5)

# your answer here

#The cities with the top five hispanic populations in 2020 were Baltimore city, Wheaton, Germantown, Silver Spring and Aspen Hill.

```



Q2: Find the cities with the largest Black population and arrange the results in descending order. 

A2:

```{r}
#write your code here

Black <- subset(Maryland, select = c(x2020_black))
summary(Black$x2020_black)

Maryland %>% 
     select(x2020_black, place) %>% 
     slice_max(x2020_black, n = 5)
# your answer here

#The five cities with the highest black population in 2020 were Baltimore city, Waldorf, Bowie, Clinton and Columbia.

```

Q3: Create a table called top_hispanic that shows the top 15 communities with the largest Hispanic population in 2020, and also display the 2010 Hispanic population and the total community population for 2020. Look at your results - what stands out to you? Craft a one-paragraph story pitch for your editor. 

A3:

```{r}

top_hispanic <- subset(Maryland, select = c(hispanic_2020))
summary(top_hispanic$hispanic_2020)


Maryland %>% 
     select(hispanic_2020, place, x2010_hispanic, x2020_total, x2010_total) %>% 
     slice_max(hispanic_2020, n = 15)



# your answer here

#Most of the hispanic populations in the top 15 increased significantly from 2010 to 2020. Frederick city, Montgomery Village and Adelphi almost doubled. More research into why this population shifted drastically could ne interesting to put together for a story. Also, another trend to note is that despite these being the top 15 for the hispanic population, the total population for the areas in 2020 was still much higher than the hispanic population. The story could also talk about the experience of hispanic people living in these areas, since they are at the same time the most populated for their identity but they are simultaneously a sizable minority compared to the rest of the population. Total population from 2010 to 2020 mostly increased by small amounts with a few exceptions. 

```
