#Lab_04 {lab_04}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here

library(tidyverse)

# Load janitor here

library(janitor)

```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.

md_race <- read_csv("lab_04.csv") %>% as.data.frame()

# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)

```

## Answer questions

Q1: Name the five cities that had the largest Hispanic population as a percentage of the entire population in 2020 and call it top5_hispanic. Produce a top 5 table for each, selecting the place, 2020 Hispanic, 2020 total and your new calculation.

A1:
```{r}

# Write your code in this codeblock.

top5_hispanic <- md_race %>% 
  select(place, x2020_total, x2020_hispanic) %>% 
  mutate(hispanic_percent = (x2020_hispanic/x2020_total)*100)



top5_hispanic %>% 
  select(x2020_hispanic, x2020_total, place, hispanic_percent) %>% 
  slice_max(x2020_hispanic, n=5)


```

Q2: Compare the top 10 cities with the largest Black population as a percentage of the overall population in 2010 and 2020. Filter for places with more than 20,000 population in 2010. Call your results pct_black. Compare results for 2020 vs 2010. What is different? What surprised you? 

A2:

```{r}
#write your code here

pct_Black_2020 <- md_race %>% 
  select(place, x2020_total, x2020_black) %>% 
  mutate(black_percent = (x2020_black/x2020_total)*100) %>% 
  slice_max(x2020_black, n=10)

pct_Black_2010 <- md_race %>% 
  select(place, x2010_total, x2010_black) %>% 
  mutate(black_percent2010 = (x2010_black/x2010_total)*100) %>% 
  slice_max(x2010_black, n=10) %>% 
  filter(x2010_black > 20000) 

#Something that stands out about these two tables is that a lot of the same cities are on the list for the highest Black populations in 2010 and 2020. For example, the top two remained Baltimore city and Waldorf, and Baltimore city's Black population decreased two percentage points while Waldorf's increased by more than 10 percentage points. Waldorf's overall population also increased by about 10,000 while Baltimore city's overall population decreased by about 2,000. I was surprised to see such differences in changes in the top two cities for the Black population, and think it could make a good story to investigate why there was such a growth in Waldorf and a loss in Baltimore city.


```

Q3: Find the 15 communities that had the great gains in Hispanic population from 2010 to 2020 as measured by percentage change. Also calculate percentage change for the community's overall growth during the same period. Filter your results for communities with more than 10,000 people as of 2020. Your results will be hisp_growth. Look at your results, craft a one-paragraph story pitch for your editor. 

A3:

```{r}

hisp_growth <- md_race %>% 
  select(place, x2010_hispanic, x2020_hispanic) %>% 
  mutate(percent_change_hispanic = (x2020_hispanic - x2010_hispanic)/x2010_hispanic) %>%
  filter(x2020_hispanic > 10000)
  
total_growth <- md_race %>% 
  select(place, x2020_total, x2010_total) %>%
  mutate(percent_change_total = (x2020_total - x2010_total)/x2010_total) %>%
  slice_max(percent_change_total, n=15)

#In the tables for the overall populations and the hispanic populations, there was no overlap between cities. I also noticed that the overall populations increased at a much faster rate than the fastest increasing hispanic populations. I think exploring why the hispanic population grew at a much slower rate and also why the total populations grew so quicly could be interesting for a story. I think also exploring what external factors members of these populations think contributed to these shifts. It would also be interesting to hear from hispanic members that moved to these communities within this time period and ask what things are important to them in a neighborhood and what drew them there.

```
