#lab_08
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse and sf

```{r}

install.packages('Rtools')

install.packages('e1071')

install.packages('sf')
install.packages('leaflet')

library(tidyverse)
library(janitor)
library(sf)
library(leaflet)

install.packages("tidycensus", install=TRUE)
library(tidycensus)

```

## Load data

```{r}

baltimore_tract_shapefiles <- read_rds("data/baltimore_tract_shapefiles.rds")
baltimore_tract_demographics <- read_rds("data/baltimore_tract_demographics.rds")
baltimore_homicides <- read_rds("data/baltimore_homicides.rds")

```

You have been given two spatial dataframes and one regular dataframe to load.  The dataframes are:

* baltimore_homicides.rds | spatial | location and details about each Baltimore homicide between 2007 and 2017, including information about the status (open, closed with arrest, closed without arrest) and the exact location. Details on the data: https://github.com/washingtonpost/data-homicides
* baltimore_tract_shapefiles.rds | spatial | polygons of each census tract in Baltimore
* baltimore_tract_demographics.rds | regular | the total population and white population of each baltimore Census tract in 2017.

**Task** Load the data

## Questions

**Q1** Are homicides that occur in majority white Census tracts in Baltimore solved at a higher rate than homicides in non-majority white Baltimore census tracts?
* Write up your findings in three paragraphs that could be dropped into a news story, with the answer to the question as the lede. 
* Include any other relevant details.
* Output a table with relevant results. 
* Output at least one map visualization pertinent to your finding.


```{r}
clean_names(baltimore_tract_demographics)

clean_names(baltimore_homicides)

baltimore_homicides <- baltimore_homicides %>%
  st_as_sf(
    coords = c("lon","lat"),
    crs="NAD83"
  )

write_rds(baltimore_homicides,"data/baltimore_homicides.rds")

census_api_key("549950d36c22ff16455fe196bbbd01d63cfbe6cf")

baltimore_tract_demographics <- get_acs(geography = "tract", state="Maryland", variables = c( "B01001_001","B02001_002"), year = 2017) %>%
  clean_names() %>%
  select(geoid, name, variable, estimate) %>%
  pivot_wider(names_from=variable, values_from=estimate) %>%
  rename(total_pop = B01001_001,
    white_pop = B02001_002
    ) %>%
  filter(str_detect(name,"Baltimore city"))

write_rds(baltimore_tract_demographics,"data/baltimore_tract_demographics.rds")


baltimore_tract_shapefiles <- get_acs(geography = "tract", state="Maryland", variables = c( "B01001_001"), year = 2017, geometry = TRUE) %>%
  clean_names() %>%
  select(geoid,name,geometry) %>%
  filter(str_detect(name,"Baltimore city"))

write_rds(baltimore_tract_shapefiles,"data/baltimore_tract_shapefiles.rds")


```

```{r}

majority_white_tracts <- baltimore_tract_demographics %>%
  mutate(pct_white = white_pop/total_pop) %>%
  select(geoid,name,pct_white) %>%
  mutate(neighborhood_type = case_when(
    pct_white > .50 ~ "majority_white",
    TRUE ~ "not_majority_white"
  ))

majority_white_tracts <- majority_white_tracts %>%
  inner_join(baltimore_tract_shapefiles) %>%
  st_as_sf()

shootings_by_neighborhood <- baltimore_homicides %>%
  st_join(majority_white_tracts) %>%
  as_tibble()

case_status_by_neighborhood <- shootings_by_neighborhood %>%
  group_by(neighborhood_type,disposition) %>%
  count() %>%
  pivot_wider(names_from=disposition, values_from=n) %>%
  clean_names() %>%
  mutate(total_shootings = closed_by_arrest+closed_without_arrest+open_no_arrest) %>%
  mutate(unsolved = open_no_arrest + closed_without_arrest) %>%
  mutate(pct_unsolved = unsolved/total_shootings*100)

case_status_by_neighborhood

ggplot() +
  geom_sf(data=baltimore_tract_shapefiles) +
  geom_sf(data=baltimore_homicides)

```

**A1** 

Majority white neighborhoods have a lower percentage of unsolved cases, according to census data. Majority white neighborhoods have an unsolved percentage of 46.4% and the not majority white neighborhoods have an unsolved rate of 66.2%. 

The amount of unsolved cases in not majority white neighborhoods is 1544.1% the amount in majority white neighborhoods. The total shootings in not majority white neighborhoods is 2,588 while the total amount of shootings in majority white neighborhoods was 239. 

The number of closed cases by arrest for not majority white neighborhoods is 682.8% the amount in majority white neighborhoods. The number of cases that are closed without arrest in not majority white neighborhoods is 913.3% the amount in majority white neighborhoods. 

The number of cases that are open with no arrest in not majority white neighborhoods is 1642.7% the amount in majority white neighborhoods. 
