#lab_08
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse

```{r}
library(janitor)
library(tidyverse)
```


## Load data

**Task**: Load four tables of U.S. Census data.  Each table has one row per Maryland county with 2020 population information.  Each table has these same two columns:

* geoid - a unique identifier for each county
* name - the county name

The third column in each table represents the total population in one of four demographic categories.

* maryland_population.csv - total_pop - total population 
* maryland_hispanic.csv - hispanic_pop - total Hispanic population
* maryland_white.csv - white_pop - total White alone population
* maryland_black.csv - black_pop - total Black alone population

Some of the information in the tables may need to be cleaned or validated in order to produce the correct answer.  

```{r}
maryland_black <- read_csv("data/maryland_black.csv")
maryland_hispanic <- read_csv("data/maryland_hispanic.csv")
maryland_population <- read_csv("data/maryland_population.csv")
maryland_white <- read_csv("data/maryland_white.csv")

glimpse(maryland_black)
glimpse(maryland_hispanic)
glimpse(maryland_population)
glimpse(maryland_white)
```


## Questions

**Q1**  What Maryland county has the largest Hispanic population (by percentage of total population)? 

**A1** 
```{r}

add_row_hispanic <- maryland_hispanic %>% 
  add_row(
    name = "Montgomery County, Maryland",
    geoid = 24031,
    hispanic_pop = 204620
  )

maryland_hispanic_total_population <- add_row_hispanic %>%
  inner_join(maryland_population, by="name", "geoid") %>% 
  mutate(percent_hispanic_of_total = (hispanic_pop/total_pop)*100) %>% 
  arrange(desc(percent_hispanic_of_total))

#The Maryland county with the largest Hispanic population by percentage of total population is Montgomery County.

```

**Q2**  What is the total population of people who identify as white alone in Maryland? 

**A2**
```{r}
  
maryland_white %>% 
  mutate(white_pop = case_when(
    name == "Somerset County, Maryland" & geoid == 24039 ~ 13510)
    )

sum(maryland_white$white_pop)

#The total population of people who identify as white alone in Maryland is 3,275,048

```

**Q3** What is the average Black population in all Maryland counties? 
**A3** 

```{r}

maryland_black %>% 
  mutate(black_pop = case_when(
    name == "Baltimore County, Maryland" &
      geoid == 24005 ~ 243098,
    name == "Baltimore city, Maryland" &
      geoid == 24510 ~ 371243, TRUE ~ black_pop
    )
    )
  
mean(maryland_black$black_pop)



#The average black population is 74973.71.

```
